Machine Learning and Constraint Programming for Relational-To-Ontology Schema Mapping

### 2017-09-03 stp/paper_aaai/main.tex
Squeezed the paper for page limit

%  \pdfinfo{
%/Title (2018 Formatting Instructions for Authors Using LaTeX)
%/Author (AAAI Press Staff)}
\setcounter{secnumdepth}{2}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Machine Learning and Constraint Programming for Relational-To-Ontology Schema Mapping }
\author{}
\maketitle
\begin{abstract}
The problem of integrating heterogeneous data sources in an ontology is of high 
relevance in the database field. Several techniques exist to approach this 
problem. Nonetheless, side constraints on the data integration cannot be easily 
integrated with these tools and thus the results may be inconsistent. We 
present a new approach that combines Machine Learning and 
Constraint Programming techniques, by modeling the data integration problem as 
a Steiner Tree Problem in a weighted graph. We show that through this approach 
we can achieve better precision, recall and speed compared to state-of-the-art 
approaches. We provide a comprehensive set of experiments supporting our 
findings.
\end{abstract}

@@

In this paper we use Machine Learning techniques (ML) to learn mapping rules from previously mapped instances.
To this end we formulate the Relational-To-Ontology Mapping Problem (\relonto{}) as a Minimum Steiner Tree Problem (STP) with side constraints \cite{deuna2016steiner}.

Firstly, we build an \emph{integration graph} which includes the 
attributes from the new source as well as ontology classes and properties. 
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use Constraint Programming (CP) to 
find a minimum cost Steiner Tree in the graph.
The goal is to assign the costs on the edges in such a way that the resulting 
Steiner Tree is a valid and coherent semantic model for the new source. 
Our contributions are:
\begin{itemize}
	\item A novel modeling framework for the \relonto{} problem which 
	seamlessly incorporates machine learning and constraint programming.
	\npr{symbolic AI, ontological reasoning, transitive closure}
	\item An efficient approach to handle attributes which cannot be matched to 
	the ontology, through a new semantic labeling model and a new STP 
	model with additional nodes.
	\item A flexible and extensible approach to the problem, which is easy to 
	reuse and adapt to specific situations.
\end{itemize}

@@

\end{tikzpicture}
\caption{Example of ontology
(`\tikz{ \node[class, inner sep = 2] () at (0,0)  {};}' are Classes, `\tikz{ \path[hasaC,->] 
(0,1) -- (3ex,1);}' means ``has an object property'', 
`\tikz{ \path[hasaA,decoration={post length=0}] (0,0) edge[decorate] (3ex,0);}' 
means ``has a data property'' and 
`\tikz{ \path[isa,->] (0,1) -- (3ex,1);}' means ``is a'' ). Example by
\citeasnoun{Taheriyan2013}.
}
\label{FIG:onto}
\end{figure}

@@

\end{scope}
\end{tikzpicture}
\caption{Examples of two semantic models (`\tikz{ \node[class, inner sep = 2] () at (0,0)  
{};}' are class nodes, and `\tikz{ \node[data] () at (0,0)  {};}' are data 
nodes).}
\label{FIG:sem}
\end{figure}

@@

We formulate the problem of semantic labeling as a multi-class classification
problem.
The known source descriptions $S_T$ provide us the training sample.
We compute a feature vector for each attribute in a data source and associate 
the known semantic type with the corresponding feature vector.
The feature vector, among others, includes characteristics
such as a number 
of whitespaces and other special characters, statistics of 
values in the column 
%(e.g, mean/max/min string length and numeric statistics)
, etc. 
A full list can be found in Appendix \ref{Ann:SemLab}.
\npr{If I remove entropy descr, we are perfectly within page limit}
%\ignore{
One of the important features characterising information content of an attribute is Shannon's entropy.
% of the attribute's concatenated rows
Shannon's entropy 
%(or information entropy~\cite{Manning:Introduction})
 of a string $X$ is defined as
$H(X) = -\sum_{i}{p_i \log_{2}p_i},$ where $p_i$ is the probability of a character, whose index in character vocabulary is $i$, to appear in $X$, and the summation ranges over all
characters in the vocabulary.
%of all possible characters
To evaluate $p_i$ in Shannon's entropy, we evaluate normalized character frequency distribution \emph{chardist} of an attribute, as character counts in concatenated rows of the attribute, normalized by the total length of the concatenated rows.
The vocabulary of all characters consists of 100 printable characters (including $\backslash$n).
Finally, we add the 100-dimensional vector of $p_i$ to the attribute feature vector.
%}
We also compute a set of features based on similarity metrics inspired by state-of-the-art works~\cite{Pham:semantic} or~\cite{Ritze:matching}.
Among others, we compute mean cosine similarity for character distributions of attribute values and string similarity metrics for attribute names.
%To extract features from attribute names, we compute string similarity metrics: minimum edit distance, two WordNet based similarity measures such as JCN~\cite{Jiang:Semantic} and LIN ~\cite{Lin:Information}, and $k$-nearest neighbors using Needle-Wunsch distance~\cite{Needleman:General}.
%The minimum edit distance between two strings $s_1$ and $s_2$ is the minimum number of edit operations, such as insertion, deletion, substitution, which are required to transform one string into another~\cite{Manning:Introduction}.
%We compute the similarity between attribute name and all class instances in the training data.
%The number of thus extracted features depends on the number of semantic labels in the training data.
We train a random forest on the obtained sample.

@@

\end{tikzpicture}
\caption{Example of alignment graph (`\tikz{ \node[class, inner sep = 2] () at (0,0)  
{};}' `\tikz{ \node[class,orange,fill=orange!30!white,text=black, inner sep = 2] () at 
(0,0)  
{};}' are class nodes, and `\tikz{ \node[data] () at (0,0)  {};}' are data 
nodes). We omit weights for clarity.}
\label{FIG:ali}
\end{figure}

@@

We associate a weighting function  $w_\mathcal{I} : E \mapsto \mathbb{R}^+$ to 
the integration graph. For an edge $e \in \mathcal{E_O}$, $w_\mathcal{I}(e) = 
w_\mathcal{O}(e)$. For an edge $e 
\in \mathcal{M}_\mathcal{O}^{s^\star}$ connecting attribute $a_i$ to the node 
$l_j$ induced by the semantic model, $w_\mathcal{I}(e) = - 
ln(\psi(a_i,l_j))$. %\ddg{CHECK!}
An example of an integration graph can be found in Fig. \ref{FIG:inte}.

@@

	\path[draw] (a4) -- (stn1);
	\path[draw] (a4) -- (cin1);
	\path[draw] (a4) -- (con1);
	\path[draw] (a4) -- (cin2);
	\path[draw] (a4) -- (con2);
\end{tikzpicture}
\caption{Example of integration graph. (`\tikz{ \node[att] () at (0,0)  
{};}' are attribute nodes). We omit weights and only show the most likely 
matches for clarity. Data node
are the same as in Fig \ref{FIG:ali}.}
\label{FIG:inte}
\end{figure}

@@

It is therefore natural to model this problem as a STP with side constraints, since the 
two first requirements imply that $T^\star$ is a Steiner Tree of the 
integration graph. By designing the weighting function $w_\mathcal{I}$ through 
ML techniques, as shown in Section \ref{SEC:ML}, 
our expectation is that the minimum weight Steiner Tree is a valid semantic 
model for the new source.

@@

As explained in Subsection \ref{SSEC:pattern-mining}, we also tried to use 
frequently appearing patterns in order to incentivise the solution tree 
$T^\star$ to contain subgraphs of the alignment graph that have been frequently 
seen in the training set.

To do this, we use the support of each of the obtained patterns as a prize. If 
the tree contains a pattern, then its weight is automatically reduced by a the 
value of the support of that pattern. We will see in the next section how this 
information is integrated in the model.

@@

% architecture
We have run all our experiments on a Dell server with 252 GB of memory, 2 CPUs 
with 4 cores each.
\npr{check characteristics, AWS machine is bigger}
In all our experiments we use a timeout threshold of 15s for \chuffed{}, which 
runs on a single core (thus easy to deploy on any user's machine).

@@

		\end{axis}
	\end{tikzpicture}
\subcaption{Recall (soccer)}\label{fig:soccerrec}
\end{minipage}%
\caption{Performance over the museum and soccer datasets. Average size of 
integration graph: 88 nodes and 1129 nodes.\label{fig:perf}}
\end{figure}

@@

We have also observed that \serenepats{} manages to find the ground truth on 
many instances among its sequential solutions, but again it takes much more 
time.

@@

\begin{axis}[
  axis y line*=right,
  axis x line=none,
  ymin=0, ymax=15,
  legend pos=south east,
  ylabel=time (s),
  		    		    style = {font=\huge}
]
\addlegendimage{/pgfplots/refstyle=plot_one}\addlegendentry{precision}
\addlegendimage{/pgfplots/refstyle=plot_two}\addlegendentry{recall}
\addplot[smooth,mark=square,blue]
  coordinates{
    (0,0.894338290115883)(1,2.39426428574782)(2,6.89703391393025)(3,10.3563708686829)(4,9.87668780326843)(5,9.92719115257263)(6,10.4045292949677)(7,14.3818371582031)(8,14.4026180458069)(9,14.8023208427429)
}; \addlegendentry{time}
\end{axis}
\end{tikzpicture}
\caption{Performance of \serene{} for leave-one-out strategy on the museum 
domain.}\label{fig:loo}
\end{figure}

>>>>

%  \pdfinfo{
%/Title (2018 Formatting Instructions for Authors Using LaTeX)
%/Author (AAAI Press Staff)}
\setcounter{secnumdepth}{2}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Machine Learning and Constraint Programming for Relational-To-Ontology Schema Mapping }
\author{}
\maketitle
\begin{abstract}
The problem of integrating heterogeneous data sources into an ontology is of high 
relevance in the database field. Several techniques exist to approach this 
problem. Nonetheless, side constraints on the data integration cannot be easily 
integrated with these tools and thus the results may be inconsistent. We 
present a new approach that combines Machine Learning and 
Constraint Programming techniques, by modeling the data integration problem as 
a Steiner Tree Problem in a weighted graph. We show that through this approach 
we can achieve better precision, recall and speed compared to state-of-the-art 
approaches. We provide a comprehensive set of experiments supporting our 
findings.
\end{abstract}

@@

In this paper we use Machine Learning techniques (ML) to learn mapping rules from previously mapped instances.
To this end we formulate the Relational-To-Ontology Mapping Problem (\relonto{}) as a Minimum cost Steiner Tree Problem (STP) with side constraints \cite{deuna2016steiner}.

Firstly, we build a graph which includes attributes from the new
source as well as ontology classes and properties.
We call it the \emph{integration graph}.
This graph uses information gathered from the previously mapped data sources, and is further extended with information derived from the ontology.
%Firstly, we build a \emph{integration graph} which includes the 
%attributes from the new source as well as ontology classes and properties.
Secondly, we apply machine learning techniques to 
assign costs to its edges. 
Lastly, we use Constraint Programming (CP) to 
find a minimum cost Steiner Tree in the graph.
The goal is to assign the costs on the edges in such a way that the resulting 
Steiner Tree is a valid and coherent semantic model for the new source. 
Our contributions are:
\begin{itemize}
	\item A novel modeling framework for the \relonto{} problem which 
	seamlessly incorporates machine learning and constraint programming.
%	\npr{symbolic AI, ontological reasoning, transitive closure}
	\item An efficient approach to handle attributes which cannot be matched to 
	the ontology, through a new semantic labeling model and a new STP 
	model with additional nodes.
	\item A flexible and extensible approach to the problem, which is easy to 
	reuse and adapt to specific situations.
\end{itemize}

@@

\end{tikzpicture}
\caption{Example of ontology
(`\tikz{ \node[class, inner sep = 2] () at (0,0)  {};}' are Classes, `\tikz{ \path[hasaC,->] 
(0,1) -- (3ex,1);}' means ``has an object property'', 
`\tikz{ \path[hasaA,decoration={post length=0}] (0,0) edge[decorate] (3ex,0);}' 
means ``has a data property'' and 
`\tikz{ \path[isa,->] (0,1) -- (3ex,1);}' means ``is a'' ). Example by
\citeasnoun{Taheriyan2013}.
}
\label{FIG:onto}
\vspace{-3mm}
\end{figure}

@@

\end{scope}
\end{tikzpicture}
\caption{Examples of two semantic models (`\tikz{ \node[class, inner sep = 2] () at (0,0)  
{};}' are class nodes, and `\tikz{ \node[data] () at (0,0)  {};}' are data 
nodes).}
\label{FIG:sem}
\vspace{-3mm}
\end{figure}

@@

We formulate the problem of semantic labeling as a multi-class classification
problem.
The known source descriptions $S_T$ provide us the training sample.
We compute a feature vector for each attribute in a data source and associate 
the known semantic type with the corresponding feature vector.
The feature vector, among others, includes characteristics
such as a number 
of whitespaces and other special characters, statistics of 
values in the column 
%(e.g, mean/max/min string length and numeric statistics)
, etc. 
A full list can be found in Appendix \ref{Ann:SemLab}.
%\npr{If I remove entropy descr, we are perfectly within page limit}
%\ignore{
One of the important features characterising information content of an attribute is Shannon's entropy.
% of the attribute's concatenated rows
Shannon's entropy 
%(or information entropy~\cite{Manning:Introduction})
 of a string $X$ is defined as
$H(X) = -\sum_{i}{p_i \log_{2}p_i},$ where $p_i$ is the probability of a character, whose index in character vocabulary is $i$, to appear in $X$, and the summation ranges over all
characters in the vocabulary.
%of all possible characters
To evaluate $p_i$ in Shannon's entropy, we evaluate normalized character frequency distribution \emph{chardist} of an attribute, as character counts in concatenated rows of the attribute, normalized by the total length of the concatenated rows.
The vocabulary of all characters consists of 100 printable characters (including $\backslash$n).
Finally, we add the 100-dimensional vector of $p_i$ to the attribute feature vector.
%}
We also compute a set of features based on similarity metrics inspired by state-of-the-art works~\cite{Pham:semantic} or~\cite{Ritze:matching}.
Among others, we compute mean cosine similarity for character distributions of attribute values and string similarity metrics for attribute names.
%To extract features from attribute names, we compute string similarity metrics: minimum edit distance, two WordNet based similarity measures such as JCN~\cite{Jiang:Semantic} and LIN ~\cite{Lin:Information}, and $k$-nearest neighbors using Needle-Wunsch distance~\cite{Needleman:General}.
%The minimum edit distance between two strings $s_1$ and $s_2$ is the minimum number of edit operations, such as insertion, deletion, substitution, which are required to transform one string into another~\cite{Manning:Introduction}.
%We compute the similarity between attribute name and all class instances in the training data.
%The number of thus extracted features depends on the number of semantic labels in the training data.
We train a random forest on the obtained sample.

@@

\end{tikzpicture}
\caption{Example of alignment graph (`\tikz{ \node[class, inner sep = 2] () at (0,0)  
{};}' `\tikz{ \node[class,orange,fill=orange!30!white,text=black, inner sep = 2] () at 
(0,0)  
{};}' are class nodes, and `\tikz{ \node[data] () at (0,0)  {};}' are data 
nodes). We omit weights for clarity.}
\label{FIG:ali}
\vspace{-3mm}
\end{figure}

@@

We associate a weighting function  $w_\mathcal{I} : E \mapsto \mathbb{R}^+$ to 
the integration graph. 
For an edge $e \in \mathcal{E_O}$, $w_\mathcal{I}(e) = 
w_\mathcal{O}(e)$. F
or an edge $e 
\in \mathcal{M}_\mathcal{O}^{s^\star}$ connecting attribute $a_i$ to the node 
$l_j$ induced by the semantic model, $w_\mathcal{I}(e) = - 
ln(\psi(a_i,l_j))$. %\ddg{CHECK!}
An example of an integration graph can be found in Fig. \ref{FIG:inte}.

@@

	\path[draw] (a4) -- (stn1);
	\path[draw] (a4) -- (cin1);
	\path[draw] (a4) -- (con1);
	\path[draw] (a4) -- (cin2);
	\path[draw] (a4) -- (con2);
\end{tikzpicture}
\caption{Example of integration graph. (`\tikz{ \node[att] () at (0,0)  
{};}' are attribute nodes). We omit weights and only show the most likely 
matches for clarity. Data node
are the same as in Fig \ref{FIG:ali}.}
\label{FIG:inte}
\vspace{-3mm}
\end{figure}

@@

It is therefore natural to model this problem as a STP with side constraints.
%, since the two first requirements imply that $T^\star$ is
%a Steiner Tree of the integration graph. 
By designing the weighting function $w_\mathcal{I}$ through 
ML techniques, as shown in Section \ref{SEC:ML}, 
our expectation is that the minimum cost Steiner Tree is a valid and coherent semantic model for the new source.

@@

As explained in Subsection \ref{SSEC:pattern-mining}, we also use 
graph patterns in order to incentivise the solution tree 
$T^\star$ to contain subgraphs of the alignment graph that have been frequently 
seen in the training set.
To do this, we use the support of each of the obtained patterns as a prize. If 
the tree contains a pattern, then its weight is automatically reduced by a the 
value of the support of that pattern. We will see in the next section how this 
information is integrated in the model.

@@

% architecture
We have run all our experiments on a Dell server with 252 GB of memory, 2 CPUs 
with 4 cores each.
%\npr{check characteristics, AWS machine is bigger}
In all our experiments we use a timeout threshold of 15s for \chuffed{}, which 
runs on a single core (thus easy to deploy on any user's machine).

@@

		\end{axis}
	\end{tikzpicture}
\subcaption{Recall (soccer)}\label{fig:soccerrec}
\end{minipage}%
\caption{Performance for the museum and soccer domains. 
Average size of integration graphs for the museum domain: 88 nodes and 1129 edges2.
Average run time for the museum domain: \serene{} 0.5s, \karma{} 0.5s, \serenepats{} 1s. \label{fig:perf}}
\vspace{-3mm}
\end{figure}

@@

We have also observed that \serenepats{} manages to find the ground truth on 
many instances among its sequential solutions.
This property of our approach can be used to configure various parameters for the model, e.g., scaling factors for pattern costs or match score.

@@

\begin{axis}[
  axis y line*=right,
  axis x line=none,
  ymin=0, ymax=15,
  legend pos=south east,
  ylabel=time (s),
  		    		    style = {font=\huge}
]
\addlegendimage{/pgfplots/refstyle=plot_one}\addlegendentry{precision}
\addlegendimage{/pgfplots/refstyle=plot_two}\addlegendentry{recall}
\addplot[smooth,mark=square,blue]
  coordinates{
    (0,0.894338290115883)(1,2.39426428574782)(2,6.89703391393025)(3,10.3563708686829)(4,9.87668780326843)(5,9.92719115257263)(6,10.4045292949677)(7,14.3818371582031)(8,14.4026180458069)(9,14.8023208427429)
}; \addlegendentry{time}
\end{axis}
\end{tikzpicture}
\caption{Performance of \serene{} for leave-one-out strategy on the museum 
domain.}\label{fig:loo}
\vspace{-3mm}
\end{figure}
